---
title: "Reproducible Data: Project 1"
output: html_document
---

This document presents the solution for the cousera Reproducible Data Couse project 1.

###Loading and preprocessing the data.

```{r,echo=TRUE}
library(plyr)
data<-read.csv(file="/users/jenny/downloads/activity.csv")

```
###Question: What is the mean total number of steps per day?
```{r count.days, echo=TRUE}
byday <- ddply(data,~date,summarise,mean.steps=mean(steps,na.rm=TRUE),total.steps=sum(steps))
hist(byday$total.steps,ylab="Total number of steps per day",main="Histogram of Data")
mean.steps<-mean(byday$total.steps,na.rm=TRUE)
median.steps<-median(byday$total.steps,na.rm=TRUE)
```
The mean number of steps per day is `r mean.steps`. The median number of steps per day is `r median.steps`.

###Question: What is the average daily activity pattern?
Here is a graph of the average number of steps taken at each interval accross the days. 
```{r computemean, echo=TRUE}
aa <- ddply(data,~interval,summarise,mean=mean(steps,na.rm=TRUE))
plot(aa$interval,aa$mean,type="l",ylab="mean number of steps for all days",xlab="5 min interval")
maxmin<-aa$interval[which(aa$mean==max(aa$mean))]
```

The maximum number of steps occurs duing the `r maxmin`th 5 minute interval. 

### Imputing missing values
```{r NAcounter, echo=TRUE}
NAcount<-length(which(is.na(data$steps)))
```
There are `r NAcount` missing values in the step count data. 

Here is a plot of the histogram of the total steps taken each day and the mean and meadian total steps for each date. The values and histogram were created by substituing all `r NAcount` missing values with the mean value for that time interval accross the remaining days.  

```{r substitute.days, echo=TRUE}
#For each NA substitute the mean for the same 5 min interval
NewData<-data
for (i in 1:nrow(aa)){
  NewData$steps[which(is.na(NewData$steps) & NewData$interval==aa$interval[[i]])]<-aa$mean[[i]]
}
byday <- ddply(NewData,~date,summarise,mean.steps=mean(steps,na.rm=TRUE),total.steps=sum(steps))
hist(byday$total.steps,ylab="Total number of steps per day",main="Histogram of Data")
mean.steps<-mean(byday$total.steps,na.rm=TRUE)
median.steps<-median(byday$total.steps,na.rm=TRUE)
```
The mean number of steps per day is `r mean.steps`. The median number of steps per day is `r median.steps`. The values are the same as they were for the unsubstituted data. Since we substituted the mean values, this makes sense. 

###Question: Are there differences in activity patterns between weekdays and weekends?
The following graph compares average weekday and weekend patterns. 
```{r weekdays, echo=TRUE}
NewData$day<-weekdays(as.Date(NewData$date))
NewData$day[which(NewData$day=="Monday")]<-"Weekday"
NewData$day[which(NewData$day=="Tuesday")]<-"Weekday"
NewData$day[which(NewData$day=="Wednesday")]<-"Weekday"
NewData$day[which(NewData$day=="Thursday")]<-"Weekday"
NewData$day[which(NewData$day=="Friday")]<-"Weekday"
NewData$day[which(NewData$day=="Saturday")]<-"Weekend"
NewData$day[which(NewData$day=="Sunday")]<-"Weekend"
NewData$day<-as.factor(NewData$day)

daysteps<-ddply(NewData,.(day,interval),summarise,mean.steps=mean(steps,na.rm=TRUE),total.steps=sum(steps))
library(lattice)
xyplot( mean.steps ~ interval| factor(day), data=daysteps,type="l",layout=c(1,2))
```


